---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Insert title of project here"
subtitle: https://github.com/sfburch/SBurchFinalProject2020
author: Samantha Burch
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
setwd("~/Desktop/Data Analytics/SBurchFinalProject2020")
getwd()

# Load your packages
library('tidyverse')
library('knitr')
library(viridis)
install.packages("rvest")
library(rvest)
install.packages("ggrepel")
library(ggrepel)

# Set your ggplot theme
mytheme <- theme_classic() +
  theme(axis.text = element_text(color = "dark blue"), 
        legend.position = "top")
theme_set(mytheme)
```

```{r, echo=FALSE}
# Load your datasets
CH4.Emissions.raw <- read.csv("./Data/Raw/CH4_emissions_raw.csv")
NDC.SDG.2.4.raw <- read.csv("./Data/Raw/ndc_sdg_2.4_raw.csv")
NDC.SDG.13.raw <- read.csv("./Data/Raw/ndc_sdg_13_raw.csv")

#Look at column names
colnames(CH4.Emissions.raw)
colnames(NDC.SDG.2.4.raw)
colnames(NDC.SDG.13.raw)

#Process the data - Select only columns of interest for both
CH4.Emissions.processed <- CH4.Emissions.raw %>%
  select(Country, Data.source, Sector, Gas, Unit, X2016)

NDC.SDG.2.4.processed <- NDC.SDG.2.4.raw %>%
  select(Country, SDG.target, Status, Sector, Climate.response)

NDC.SDG.13.processed <- NDC.SDG.13.raw %>%
  select(Country, SDG.target, Status, Sector, Climate.response)

#Save the new processed dataset
write.csv(CH4.Emissions.processed, row.names = FALSE, file = "./Data/Processed/CH4_emissions_processed.csv")

write.csv(NDC.SDG.2.4.processed, row.names = FALSE, file = "./Data/Processed/ndc_sdg_2.4_processed.csv")

write.csv(NDC.SDG.13.processed, row.names = FALSE, file = "./Data/Processed/ndc_sdg_13.processed.csv")

#Combine datasets (using 'inner_join')
Combined.CH4.NDC.SDG <- inner_join(CH4.Emissions.processed, NDC.SDG.2.4.processed, NDC.SDG.13.processed)

```


# Rationale and Research Questions



\newpage

# Dataset Information
```{r}

# Geom Bar Plot
CH4GeomPlot <- ggplot(data, aes(x=Country, y=MethaneEmissions.MtCOCCe)) + 
  geom_bar(stat = "identity", width=0.2, fill="#FF9999", colour="black") + 
    theme(plot.title = element_text(face = "bold", size = 15, color = "deepskyblue4", hjust = 0.5), plot.subtitle = element_text(face = "italic", size = 13, hjust = 0.5), axis.text.x = element_text(size=12, angle=45, hjust=1, vjust=1),  axis.title.x = element_text(face = "italic", size=13), axis.title.y = element_text(face = "italic", size=13)) +
    scale_fill_brewer(palette = "Set1") +
  labs(title="Top 20 Methane (CH4) Emitters", subtitle = "Agricultural Sector, By Country (2016)", x="Country",y="CH4 Emissions (MtCO₂e)")

print(CH4GeomPlot)

```

\newpage

# Exploratory Analysis 
what they need or want before actual analysis; distributions of variables of interest; time series sampling over time; potential relationships between variables; here's the intro to the data set, might not be statsistical but help someone orient themselves, provide rationale


\newpage

# Analysis


## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
