---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Insert title of project here"
subtitle: https://github.com/sfburch/SBurchFinalProject2020
author: Samantha Burch
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
setwd("~/Desktop/Data Analytics/SBurchFinalProject2020")
getwd()

# Load your packages
library('tidyverse')
library('knitr')
library(viridis)
install.packages("rvest")
library(rvest)
install.packages("ggrepel")
library(ggrepel)

# Set your ggplot theme
mytheme <- theme_classic() +
  theme(axis.text = element_text(color = "dark blue"), 
        legend.position = "top")
theme_set(mytheme)
```

```{r, echo=FALSE}
# Load your datasets
CH4.Emissions.raw <- read.csv("./Data/Raw/CH4_emissions_raw.csv")
NDC.SDG.raw <- read.csv("./Data/Raw/ndc_sdg_raw.csv")

#Look at column names
colnames(CH4.Emissions.raw)
colnames(NDC.SDG.raw)

#Process the data - Select only columns of interest for both

#Combine datasets

#Save the new process dataset
```


# Rationale and Research Questions



\newpage

# Dataset Information



\newpage

# Exploratory Analysis 
what they need or want before actual analysis; distributions of variables of interest; time series sampling over time; potential relationships between variables; here's the intro to the data set, might not be statsistical but help someone orient themselves, provide rationale


\newpage

# Analysis


## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
