---
title: "CAIT Data Wrangling for the Final Project - SBurch 2020"
output: html_document
---

```{r setup, include=FALSE}
# Data Wrangling

```{r setup, include=FALSE}
# Set your working directory
setwd("~/Desktop/Data Analytics/SBurchFinalProject2020")
getwd()

# Load your packages
library('tidyverse')
library('knitr')
library(viridis)
install.packages("rvest")
library(rvest)
install.packages("ggrepel")
library(ggrepel)
install.packages("unpivotr")
library(unpivotr)
install.packages("ggmap")
library(ggmap)
install.packages("dplyr")
library(dplyr)
install.packages("rnaturalearth")
library(rnaturalearth)
install.packages("rnaturalearthdata")
library(rnaturalearthdata)
install.packages("sf")
library(sf)
install.packages("rgeos")
library(rgeos)
install.packages("sp")
library(sp)
install.packages("maps")
library(maps)


# Set your ggplot theme
mytheme <- theme_classic() +
  theme(axis.text = element_text(color = "dark blue"), 
        legend.position = "top")
theme_set(mytheme)
```

```{r, echo=FALSE}
# Load your datasets
CH4.Emissions.raw <- read.csv("../Data/Raw/CH4_emissions_raw.csv")
HistoricalTop20CH4.Emissions.raw <- read.csv("../Data/Raw/CAIT_Historical_CH4_emissions_raw.csv")
NDC.SDG.2.4.raw <- read.csv("../Data/Raw/ndc_sdg_2.4_raw.csv")
NDC.SDG.13.raw <- read.csv("../Data/Raw/ndc_sdg_13_raw.csv")

#Look at column names
colnames(CH4.Emissions.raw)
print(colnames(CH4.Emissions.raw))
colnames(NDC.SDG.2.4.raw)
colnames(NDC.SDG.13.raw)

#Process the data - Select only columns of interest for both

CAIT.CH4.Emissions.processed <- CH4.Emissions.raw %>%
  select(Country, Data.source, Sector, Gas, Unit, X2016)

colnames(CAIT.CH4.Emissions.processed)

CAIT.NDC.SDG.2.4.processed <- NDC.SDG.2.4.raw %>%
  select(Country, SDG.target, Indc.text, Status, Sector, Climate.response)

CAIT.NDC.SDG.13.processed <- NDC.SDG.13.raw %>%
  select(Country, SDG.target, Indc.text, Status, Sector, Climate.response)

#Save the new processed dataset
write.csv(CAIT.CH4.Emissions.processed, row.names = FALSE, file = "../Data/Processed/CH4_emissions_processed.csv")

write.csv(CAIT.NDC.SDG.2.4.processed, row.names = FALSE, file = "../Data/Processed/ndc_sdg_2.4_processed.csv")

write.csv(CAIT.NDC.SDG.13.processed, row.names = FALSE, file = "../Data/Processed/ndc_sdg_13.processed.csv")

#Combine datasets
Combined.CH4.2.4 <- full_join(CAIT.CH4.Emissions.processed, CAIT.NDC.SDG.2.4.processed)

Combined.CH4.13 <- full_join(CAIT.CH4.Emissions.processed, CAIT.NDC.SDG.13.processed)

Combined.2.4.13 <- rbind(CAIT.NDC.SDG.2.4.processed, CAIT.NDC.SDG.13.processed)

ColRemovedCombined.2.4.13 <- Combined.2.4.13 %>%
  select(Country, SDG.target)

FinalJoinCH4SDGs <- full_join(CAIT.CH4.Emissions.processed, ColRemovedCombined.2.4.13) %>%
  group_by(Country, X2016) %>%
  summarise(Num_Target=n()) %>%
  ungroup() %>%
  filter(Country %in% c("India", "China", "Brazil", "European Union", "United States", "Pakistan", "Indonesia", "Argentina", "Australia","Ethiopia", "Myanmar", "Mexico", "Bangladesh", "Russia", "Nigeria", "Vietnam", "Philippines", "Thailand", "Sudan", "Chad", "France")) %>%
  mutate(Country=ifelse(Country=="United States", "United States of America", Country))

write.csv(FinalJoinCH4SDGs, row.names = FALSE, file = "../Data/Processed/FinalJoinCH4SDGs_Processed.csv")

```
